---
title: "Practical ML Course Project"
author: "Rishab Ravi"
output: html_document
---
  
#Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively.  
These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.  
One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. This project uses data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.  
They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).  


```{r warning=FALSE, message=FALSE}
#loading libraries
library(caret)
library(randomForest)
```


```{r warning=FALSE, message=FALSE}
#reading data
trainData <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!",""))
testData <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
```

##Data preprocessing
```{r}
#remove nzv
subTrain <- trainData[, names(trainData)[!(nzv(trainData, saveMetrics = T)[, 4])]]
#remove NA cols
subTrain <- subTrain[, names(subTrain)[sapply(subTrain, 
          function (x)  ! (any(is.na(x) | x == "")))]]

#remove unlikely predictor variables
subTrain <- subTrain[,-1]
subTrain <- subTrain[, c(1:3, 5:58)]

```

```{r}
#Divide the training data into sub-training/test set
set.seed(3213)
inTrain <- createDataPartition(subTrain$classe, p = 0.6, list = FALSE)
subTraining <- subTrain[inTrain,]
subTest <- subTrain[-inTrain,]
```

##Model Building  

##1. Random Forest  
```{r}
set.seed(12345)
#controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFit1 <- train(classe ~ ., data=subTraining, method="rf")
pred1 <- predict(modFit1, subTraining)
confusionMatrix(pred1, subTraining$classe)
```

The accuracy is **100%!**.  

##2. knn
```{r}
set.seed(100)
ctrlKNN <- trainControl(method="adaptive_cv")
modFit2 <- train(classe~.,data=subTraining,method="knn",trControl = ctrlKNN)
pred2 <- predict(modFit2, newdata = subTraining)
confusionMatrix(pred2,subTraining$classe)
```

The accuracy is 43.64%.  

##3. GBM
```{r}
set.seed(10021)
ctrlGBM <- trainControl(method = "adaptive_cv")
modFit3 <- train(classe ~ ., data=subTraining, method = "gbm", trControl = ctrlGBM, verbose = FALSE)
gbmFinMod <- modFit3$finalModel
pred3 <- predict(modFit3,newdata=subTraining)
confusionMatrix(pred3,subTraining$classe)
```

The accuracy of our GBM model is **99.78%**.  

The **random forest** model is the better predictor, and we shall use the same to apply to the testing dataset.  

```{r}
#details of model:
varImp(modFit1) #important predictor variables

modFit1$finalModel # final model
```

##Test data prediction
```{r}
#Apply to test data set
predTesting <- predict(modFit1, testData)
predTesting
```

